\Title

### 2022-11-30 Grimaldi-etal-BiolCybernetics.tex
Update on Overleaf.

\subsection{Task definition: fast motion detection}
%%%-----------------------------------------------------------------
Let us now define a procedure of animating of natural scene by virtual eye-movements  similar to that which was used in neurobiological~\citep{vinje_sparse_2000, baudot_animation_2013} and computational neuroscience~\citep{kremkow_push-pull_2016} studies. First, let's define a trajectory inspired by the biological movements of the eyes. Indeed, these movements allow us to dynamically actuate the center of vision, or gaze, in the field of vision. In animals with a fovea, this is particularly useful as it allows to move the area with the highest density of photoreceptors in the environment, for example at a point of interest. A specific mechanism to do so are called saccades which are rapide eye movements of the eyeball that reposition the center of vision. In humans, these are very frequent (on average $2$ each second~\citep{dandekar_neural_2012}). They are produced very rapidly (about $80~\ms$) and display a whole range of speeds. At a more microscopic scale, the human gaze moves with an incessant drift similar to a Brownian-like trajectory~\citep{poletti_head-eye_2015}. To maintain the full generality of the task, we will define eye movements using a form of random walk~\citep{engbert_integrated_2011}. This approach first defines a finite set of possible 2D motions in polar coordinates. Based on the distribution of biological motions, we simplified it by selecting a set of eye movements as the Cartesian product of % N-V-phi
$8$ linearly spaced motion directions and %N-Y-velocity=
$7$ different velocities (see inset of \seeFig{motion_task}-(a)). Note that the velocities are sampled on a geometric scale between $1/4$ and $4$. Next, we define one trajectory of gaze as segments whose duration is drawn from a Poisson distribution with an average block length of $20~\ms$, similarly to a Lévy flight~\citep[p. 289]{mandelbrot_fractal_1982}. Finally, the trajectory is integrated by assuming first that velocities are uniformly and independently sampled from the set of different motion-sets and second that motion is uniform during a time segment. The resulting instances yield trajectories qualitatively similar to those observed for human eye movements (\seeFig{motion_task}-(a)).

@@

% \note{ TODO: formula / say it is driven by image gradients / give scaling rules }
To transform each movie into events, we compute a gradient image (initialized at zero) by adding the gradient of the pixels' intensity over two successive frames. If, on a specific pixel at that specific timestamp, the absolute value of this gradient exceeds a threshold, an event is generated. The event has either an OFF or ON polarity, respectively whether the gradient is negative or positive. This signed threshold value is then subtracted from the residual gradient image. When applied to the whole movie, the event stream is as a consequence similar to the output of a neuromorphic camera~\citep{rasetto_challenges_2022}, that is, a list of events defined by $x_\arank$ and $y_\arank$ (their position on the pixel grid), their polarity $\polev_\arank$ (ON or OFF) and time $\timev_\arank$  (\seeFig{motion_task}-(c)). The goal here is to infer the correct motion solely by observing these events. 
This sensory signal representing the output of this event-based camera forms a discrete stream of events, $\event$ as defined above.   
Each event has an associated address, which is typically in the form $\presynaddr_\arank=(x_\arank, y_\arank, \polev_\arank)$. This defines a presynaptic address space $\presynaddrspace = [1, \Nx]\times[1, \Ny] \times [1, \Npol] \subset \mathbb{N}^3$ where $(\Nx, \Ny)$ is the size of the sensor in pixels and $\Npol$ is the number of polarities  ($\Npol=2$ for ON and OFF polarities). 
% \note{ the task is full-field rigid translations, the model resolves it locally (analogy to the input from V1 to MT) }
%
\subsection{Extension as 3D convolutions for motion detection}
%%%-----------------------------------------------------------------
%
Now that our motion detection task is defined in visual space, we can extend the convolution-based detection model to a 3D convolution so that the resulting model would also benefit from spatial invariance. The use of spatio-temporal filters on an event stream has improved the performance of CNNs for an action recognition task in a model by~\citet{ghosh_spatiotemporal_2019} and here we will use a similar strategy, but focus on using an event-based representation.
For this purpose, we design 3D kernels of shape $(\Kx, \Ky, \Ktime) = (9, 9, 11)$, representing the two spatial dimensions and the range of time delays, respectively.
The computations are performed on the spatio-temporal windows that are defined by the kernels and are ``moving'' around the events, i.e. the spatial center of the spatio-temporal window around the current event $\epsilon_\arank$. Note that to remain within the framework of a causal calculation, the kernels are shifted in time such that only past information gives an answer at the present time.
%
The output of the MLR model corresponds to an event with the highest probability class, keeping the same timing as the input event. It is important to note that we include the prior knowledge that the event-based input is spatio-temporally clustered and that these clusters are sparse. Therefore, we will only select a given proportion of the most active cells, which we set here to $.01\%$ of all output voxels. This threshold was found using validation and future work should automate how this may be determined for various datasets. Given a stimulus input, the model thus yields a spike output in the postsynaptic address space.

%
Since the model is fully differentiable, we can now implement a semi-supervised learning rule. The loss function of the MLR model is the binary cross-entropy on the output of the classification layer and can be extended to space-time convolutions. Supervision was implemented using the input binary events as defined above and the labels as the desired output. The labels were defined at each time point as a one-shot encoding of the current motion in the channel corresponding to the current motion for all positions. As in this semi-supervised context, the label is known, but the timing is not, we used a selection of the temporal support in an unsupervised manner. It is important to note that we weighted the cost function by considering only active cells, such that the error is only back propagated to the spatial locations of these most active cells. This is reminiscent of previous methods solving this problem using a Winner-Takes-All mechanism~\citep{masquelier_unsupervised_2007}. Simulations are performed with the PyTorch library using gradient descent with Adam (for $2^{12}$ movies and a learning rate of $10^{-5}$). We tested the effect of different parameters on a validation set to quantify the role of each parameter.

@@

We also observe that these cells show an orientation selectivity, similar to that observe in MT neurons~\citep{deangelis_functional_1999}. Interestingly, the relative organization of the receptive fields in quadrature of phase follows a push-pull organization predicted by~\citet{kremkow_push-pull_2016} to explain neurophysiological results~\citep{baudot_animation_2013}.  

>>>>

\subsection{Task definition: fast motion detection}
%%%-----------------------------------------------------------------
Let us now define a procedure of animating of natural scene by virtual eye-movements  similar to that which was used in neurobiological~\citep{vinje_sparse_2000, baudot_animation_2013} and computational neuroscience~\citep{kremkow_push-pull_2016} studies. First, let's define a trajectory inspired by the biological movements of the eyes. Indeed, these movements allow us to dynamically actuate the center of vision, or gaze, in the field of vision. In animals with a fovea, this is particularly useful as it allows to move the area with the highest density of photoreceptors in the environment, for example at a point of interest. A specific mechanism to do so are called saccades which are rapide eye movements of the eyeball that reposition the center of vision. In humans, these are very frequent (on average $2$ each second~\citep{dandekar_neural_2012}). They are produced very rapidly (about $80~\ms$) and display a whole range of speeds. At a more microscopic scale, the human gaze moves with an incessant drift similar to a Brownian-like trajectory~\citep{poletti_head-eye_2015}. To maintain the full generality of the task, we will define eye movements using a form of random walk~\citep{engbert_integrated_2011}. This approach first defines a finite set of possible 2D motions in polar coordinates. Based on the distribution of biological motions, we simplified it by selecting a set of eye movements as the Cartesian product of % N-V-phi
$8$ linearly spaced motion directions and %N-Y-velocity=
$7$ different velocities (\seeFig{motion_task}-(a)-inset). Note that the velocities are sampled on a geometric scale between $1/4$ and $4$. Next, we define one trajectory of gaze as segments whose duration is drawn from a Poisson distribution with an average block length of $20~\ms$, similarly to a Lévy flight~\citep[p. 289]{mandelbrot_fractal_1982}. Finally, the trajectory is integrated by assuming first that velocities are uniformly and independently sampled from the set of different motion-sets and second that motion is uniform during a time segment. The resulting instances yield trajectories qualitatively similar to those observed for human eye movements (\seeFig{motion_task}-(a)).

@@

% \note{ TODO: formula / say it is driven by image gradients / give scaling rules }
To transform each movie into events, we compute a gradient image (initialized at zero) by adding the gradient of the pixels' intensity over two successive frames. If, on a specific pixel at that specific timestamp, the absolute value of this gradient exceeds a threshold, an event is generated. The event has either an OFF or ON polarity, respectively whether the gradient is negative or positive. This signed threshold value is then subtracted from the residual gradient image. When applied to the whole movie, the event stream is as a consequence similar to the output of a neuromorphic camera~\citep{rasetto_challenges_2022}, that is, a list of events defined by $x_\arank$ and $y_\arank$ (their position on the pixel grid), their polarity $\polev_\arank$ (ON or OFF) and time $\timev_\arank$  (\seeFig{motion_task}-(c)). The goal here is to infer the correct motion solely by observing these events. 
This sensory signal representing the output of this event-based camera forms a discrete stream of events, $\event$ as defined above.   
Each event has an associated address, which is typically in the form $\presynaddr_\arank=(x_\arank, y_\arank, \polev_\arank)$. This defines a presynaptic address space $\presynaddrspace = [1, \Nx]\times[1, \Ny] \times [1, \Npol] \subset \mathbb{N}^3$ where $(\Nx, \Ny)$ is the size of the sensor in pixels and $\Npol$ is the number of polarities  ($\Npol=2$ for ON and OFF polarities). 
% \note{ the task is full-field rigid translations, the model resolves it locally (analogy to the input from V1 to MT) }
%
\subsection{Extension as 3D convolutions for motion detection}
%%%-----------------------------------------------------------------
%
Now that our motion detection task is defined in visual space, we can extend the convolution-based detection model to a 3D convolution so that the resulting model would also benefit from spatial invariance. The use of spatio-temporal filters on an event stream has improved the performance of CNNs for an action recognition task in a model by~\citet{ghosh_spatiotemporal_2019} and here we will use a similar strategy, but focus on using an event-based representation.
For this purpose, we design 3D kernels of shape $(\Kx, \Ky, \Ktime) = (9, 9, 11)$, representing the two spatial dimensions and the range of time delays, respectively.
The computations are performed on the spatio-temporal windows that are defined by the kernels and are ``moving'' around the events, i.e. the spatial center of the spatio-temporal window around the current event $\epsilon_\arank$. Note that to remain within the framework of a causal calculation, the kernels are shifted in time such that only past information gives an answer at the present time.
%
The output of the MLR model corresponds to an event with the highest probability class, keeping the same timing as the input event. It is important to note that we include the prior knowledge that the event-based input is spatio-temporally clustered and that these clusters are sparse. Therefore, we will only select a given proportion of the most active cells, which we set here to $.01\%$ of all output voxels. This threshold was found using cross-validation and future work should automate how this may be determined for various datasets. Given a stimulus input, the model thus yields a spike output in the postsynaptic address space.

%
Since the model is fully differentiable, we can now implement a semi-supervised learning rule. The loss function of the MLR model is the binary cross-entropy on the output of the classification layer that can be extended to space-time convolutions. Supervision was implemented using the input binary events as defined above and the labels as the desired output. The labels were defined at each time point as a one-shot encoding of the current motion in the channel corresponding to the current motion for all positions. As in this semi-supervised context, the label is known, but the timing is not, we used a selection of the temporal support in an unsupervised manner. It is important to note that we weighted the cost function by considering only active cells, such that the error is only back propagated to the spatial locations of these most active cells. This is reminiscent of previous methods solving this problem using a Winner-Takes-All mechanism~\citep{masquelier_unsupervised_2007}. Simulations are performed with the PyTorch library using gradient descent with Adam (for $2^{12}$ movies and a learning rate of $10^{-5}$). We tested the effect of different parameters on a validation set to quantify the role of each parameter.
Since the model is fully differentiable, we can now implement a semi-supervised learning rule. The loss function of the MLR model is the binary cross-entropy on the output of the classification layer and can be extended to space-time convolutions. Supervision was implemented using the input binary events as defined above and the labels as the desired output. The labels were defined at each time point as a one-shot encoding of the current motion in the channel corresponding to the current motion for all positions. As in this semi-supervised context, the label is known, but the timing is not, we used a selection of the temporal support in an unsupervised manner. It is important to note that we weighted the cost function by considering only the  most active cells, such that the error is only back propagated to the spatial locations of these most active cells. This is reminiscent of previous methods solving this problem using a Winner-Takes-All mechanism~\citep{masquelier_unsupervised_2007}. Simulations are performed with the PyTorch library using gradient descent with Adam (for $2^{12}$ movies and a learning rate of $10^{-5}$). We tested the effect of different parameters on a validation set to quantify the role of each parameter.

@@

We also observe that these cells show an orientation selectivity, similar to that observed in MT neurons~\citep{deangelis_functional_1999}. Interestingly, the relative organization of the receptive fields in quadrature of phase follows a push-pull organization predicted by~\citet{kremkow_push-pull_2016} to explain neurophysiological results~\citep{baudot_animation_2013}.  
